---
title: "generate metadata"
output: github_document
editor_options:
  chunk_output_type: console
---

We will use the csvy package to add metadata to our stream chemistry data.

```{r libraries, message=FALSE}

library(csvy)
library(tidyverse)
```

## Iris data demo

First, let us see how this works with the `iris` data.

```{r iris_data_demo}

# create our data set
iris_subset <- head(iris) # we can use a small piece of iris for this demo

# add attributes
attr(iris_subset, "title") <- "the famous Iris data set" # add a title
attr(iris_subset, "description") <- "measurements of Iris (Iris spp.) plants" # add a description
attr(iris_subset, "format") <- "csvy" # add a (meta)data format
```


How does our iris data set look (in R)?

* with the str() function...

```{r iris_subset structure}

str(iris_subset)
```

*  with the attributes() function (note that we are exluding `row.names` from the output)...

```{r iris_subset attributes}

attributes(iris_subset)[!names(attributes(iris_subset)) %in% c("row.names")]
```

How does our iris data set look outside of R?

```{r write_iris_csvy}

write_csvy(x = iris_subset,
           file = '~/Desktop/iris_subset.csv',
           comment_header = F)
```


## Add metadata to our stream chemistry data

Import our stream chemistry data

```{r import_chem_data, warning=FALSE}

stream_chem <- read_csv(file = '/home/srearl/localRepos/rdm-lecture-metadata/data/stream_chemistry_metabolism.csv') %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) # fix date field on import
```


Add metadata to our stream chemistry data. We can add table- and field-level
metadata. We can loosely follow the Frictionless Data Table schema:
https://frictionlessdata.io/specs/table-schema/

```{r stream_chem_metadata}

# add table attributes
attr(stream_chem, "title") <- "stream chemistry data from southern Appalachian mountain streams, 2013" # add a title
attr(stream_chem, "description") <- "stream chemistry data for the calculation of ecosystem metabolism collected with automated data loggers from four headwater streams in the southern Appalachian mountain region during summer 2013" # add a description
attr(stream_chem, "format") <- "csvy" # add a (meta)data format

# add column attributes
attr(stream_chem$site_id, "description") <- "site-code: Greenbriar (GB), Stone Crop (SC), Hugh White Creek (HWC)"
attr(stream_chem$K2_20, "description") <- "oxygen reaeration coefficient (per day)"
attr(stream_chem$Date, "description") <- "date of observation (YYYY-MM-DD)"
attr(stream_chem$Time, "description") <- "time of observation (hh:mm:ss)"
attr(stream_chem$Temp, "description") <- "stream water temperature (degrees Celsius)"
attr(stream_chem$SpCond, "description") <- "stream water specific conductance (microSiemens per centimeter)"
attr(stream_chem$DO, "description") <- "stream water dissolved oxygen concentration (milligrams per liter)"
```


Review our stream chemistry metadata

* with the str() function...

```{r stream_chem str}

str(stream_chem)
```

*  with the attributes() function (note that we are exluding `row.names` from the output)...

```{r stream_chem attributes}

attributes(stream_chem)[!names(attributes(stream_chem)) %in% c("row.names")]
```

Write our stream chemistry data as a csvy file with metadata.

```{r write_stream_chem}

write_csvy(x = stream_chem, 
           file = "~/Desktop/stream_chem.csv",
           comment_header = F,
           row.names = F)
```

