---
title: "generate metadata"
output:
  github_document:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r libraries-for-csvy, message=FALSE, echo=FALSE}

library(csvy)
library(tidyverse)
```

## I. add metadata to tabular data as yaml with csvy

#### a simple demonstration with the iris data set

First, let us see how this works with the `iris` data.

```{r subset iris data}

# create our data set
iris_subset <- head(iris) # we can use a small piece of iris for this demo
```

We can add table-level metadata in the form of attributes to the iris_subset. As
much as possible, we will employ the Frictionless [Data Table
Schema](https://frictionlessdata.io/specs/table-schema/).

```{r iris_subset table metadata}

attr(iris_subset, "title") <- "the famous Iris data set" # add a title
attr(iris_subset, "description") <- "measurements of Iris (Iris spp.) plants" # add a description
attr(iris_subset, "format") <- "csvy" # add a (meta)data format
```

We can also add field-level metadata, including a title and description using
attributes. Table Schema does not support units metadata for fields but we can
include this important metadata in the description.

```{r iris_subset field metadata}

attr(iris_subset$Petal.Length, "label") <- "the length of an individual petal" # label will ultimately map to title
attr(iris_subset$Petal.Length, "description") <- "the length of an individual petal in units of millimeter" # add a description
```

How does our iris data set look (in R)?

* with the str() function...

```{r iris_subset structure}

str(iris_subset)
```

*  with the attributes() function (note that we are exluding `row.names` from the output)...

```{r iris_subset attributes}

attributes(iris_subset)[!names(attributes(iris_subset)) %in% c("row.names")]
```

How does our iris data set look outside of R?

```{r write_iris_csvy}

write_csvy(x = iris_subset,
           file = '~/Desktop/iris_subset.csv',
           comment_header = F)
```


#### a richer example: add metadata to our stream chemistry data

Use the csvy package to add metadata to our stream chemistry data.

Import our stream chemistry data

```{r import-chem-data, warning=FALSE}

stream_chem <- read_csv(file = 'stream_chemistry_metabolism.csv')
```

We can use the same approach we used with the subset of iris data for our stream
chemsitry data.

Add table-level metadata to our stream chemistry data.

```{r stream_chem_metadata}

attr(stream_chem, "title") <- "stream chemistry data from southern Appalachian mountain streams, 2013" # add a title
attr(stream_chem, "description") <- "stream chemistry data for the calculation of ecosystem metabolism collected with automated data loggers from four headwater streams in the southern Appalachian mountain region during summer 2013" # add a description
attr(stream_chem, "format") <- "csvy" # add a (meta)data format
```

Add table-level metadata to our stream chemistry data.

```{r}

attr(stream_chem$site_id, "label") <- "study site identifier"
attr(stream_chem$site_id, "description") <- "site-code: Greenbriar (GB), Stone Crop (SC), Hugh White Creek (HWC)"

attr(stream_chem$K2_20, "label") <- "a metric of oxygen exchange across the air-water interface"
attr(stream_chem$K2_20, "description") <- "oxygen reaeration coefficient (per day)"

attr(stream_chem$Date, "label") <- "date of observation"
attr(stream_chem$Date, "description") <- "date of observation (YYYY-MM-DD)"

attr(stream_chem$Time, "label") <- "time of observation"
attr(stream_chem$Time, "description") <- "time of observation (hh:mm:ss)"

attr(stream_chem$Temp, "label") <- "stream water temperature"
attr(stream_chem$Temp, "description") <- "stream water temperature (degrees Celsius)"

attr(stream_chem$SpCond, "label") <- "stream water specific conductance"
attr(stream_chem$SpCond, "description") <- "stream water specific conductance (microSiemens per centimeter)"

attr(stream_chem$DO, "label") <- "stream water dissolved oxygen concentration"
attr(stream_chem$DO, "description") <- "stream water dissolved oxygen concentration (milligrams per liter)"
```

Review our stream chemistry metadata

* with the str() function...

```{r stream_chem str}

str(stream_chem)
```

*  with the attributes() function (note that we are exluding `row.names` from the output)...

```{r stream_chem attributes}

attributes(stream_chem)[!names(attributes(stream_chem)) %in% c("row.names")]
```

Write our stream chemistry data as a csvy file with metadata.

```{r write_stream_chem}

write_csvy(x = stream_chem, 
           file = "~/Desktop/stream_chem.csv",
           comment_header = F,
           row.names = F)
```

## II. frictionless

```{r libraries-for-frictionless, message=FALSE, echo=FALSE}

library(tableschema.r)
library(datapackage.r)
library(future)
library(jsonlite)
```

### IIa. use the frictionless web-based data package creator

We can use the frictionless data package
[creator](https://create.frictionlessdata.io/) to construct an entire data
package. Metadata are contained in the `stream-chem-data-package.json` file in
the `data` directory of this repository; this can be uploaded along with the
data file `stream_chemistry_metabolism.csv` or bundled as a stand-along data
package.

### IIb. use frictionless tools to generate metadata or create a data package

Load an existing schema into our R environment and display output

```{r load-display-steam-chem-json, echo=FALSE}

# note that the full path to the file is required for Package.load

streamChemPackage <- Package.load('/home/srearl/localRepos/rdm-lecture-metadata/data/stream-chem-data-package.json')
toJSON(streamChemPackage$descriptor, pretty = T)
```


## III. include metadata in a well-constructed README

[Wall et al. 2017](https://datadryad.org/stash/dataset/doi:10.5061/dryad.5vg70)