---
title: "generate metadata"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

We will use the csvy package to add metadata to our stream chemistry data.

```{r libraries, echo=FALSE}

library(csvy)
library(tidyverse)
```

# iris data demo

First, let us see how this works with the iris data.

```{r iris_data_demo}

# create our data set
iris_subset <- head(iris) # we can use a small piece of iris for this demo

# add attributes
attr(iris_subset, "title") <- "the famous Iris data set" # add a title
attr(iris_subset, "description") <- "stats about Iris (Iris spp.) plants" # add a description
attr(iris_subset, "format") <- "csvy" # add a (meta)data format
```


How does our iris data set look (in R)?

```{r iris_subset structure}

str(iris_subset)
```


How does our iris data set look (outside of R)?

```{r write_iris_csvy, echo=FALSE}

write_csvy(iris_subset, '~/Desktop/iris_subset.csv', comment_header = F)
```


# Add metadata to our stream chemistry data

Import our stream chemistry data

```{r import_chem_data, echo=FALSE}

stream_chem <- read_csv('stream_chemistry_metabolism.csv') %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) # fix date field on import
```


Add metadata to our stream chemistry data. We can add table- and field-level attribute data. We can loosely follow the Frictionless Data Table schema: https://frictionlessdata.io/specs/table-schema/

```{r stream_chem_metadata}

# add table attributes
attr(stream_chem, "title") <- "stream chemistry data from southern Appalachian mountain streams, 2013" # add a title
attr(stream_chem, "description") <- "stream chemistry data for the calculation of ecosystem metabolism collected with automated data loggers from four headwater streams in the southern Appalachian mountain region during summer 2013" # add a description
attr(stream_chem, "format") <- "csvy" # add a (meta)data format

# add column attributes
attr(stream_chem$site_id, "description") <- "site-code: Greenbriar (GB), Stone Crop (SC), Hugh White Creek (HWC)"
attr(stream_chem$K2_20, "description") <- "oxygen reaeration coefficient (per day)"
attr(stream_chem$Date, "description") <- "date of observation (YYYY-MM-DD)"
attr(stream_chem$Time, "description") <- "time of observation (hh:mm:ss)"
attr(stream_chem$Temp, "description") <- "stream water temperature (degrees Celsius)"
attr(stream_chem$SpCond, "description") <- "stream water specific conductance (microSiemens per centimeter)"
attr(stream_chem$DO, "description") <- "stream water dissolved oxygen concentration (milligrams per liter)"
```


Review our stream chemistry metadata

```{r review_stream_chem_metadata}

str(stream_chem)
attributes(stream_chem)
```


Write our stream chemistry data as a csvy file with metadata.

```{r write_stream_chem, echo=FALSE}

write_csvy(stream_chem, "~/Desktop/stream_chem.csv", comment_header = F, row.names = F)
```

